---
title: "R Basics"
author: "Drew Schmidt"
date: "April 2, 2015"
output:
  ioslides_presentation:
    highlight: zenburn
    logo: pics/utk.png
    transition: faster
    smaller: no
    widescreen: yes
    css: style.css
---

## Contents




# Getting Started

## 
```{r, error=TRUE}
2+2
a <- 2+2
a
A
```

##
```{r}
c(1, 2, 3, 4, 5)
1:5
```

## 
```{r}
x <- 1:100
sum(x)
mean(x)
sd(x)
summary(x)
```

## Inspecting Objects
```{r}
head(x)
tail(x)
str(x)
```

## Vectorization
```{r}
head(x+10)
head(x*x)
```

## Getting Help
```{r}
?mean
example(mean)
?clustering
??clustering
```

## File Paths and Data Input
```r
setwd("/home/user/")
setwd("C:/Documents and Settings/user")

mydata <- read.csv("mydata.csv")
summary(mydata)
```

## Writing Data
```r
write.csv(mydata, file="mydata.csv")

saveRDS(object=mydata, file="mydata.rds")

### load rds via load()
load("mydata.rds")
mydata
```

## Addon Packages
```r
install.packages("devtools")
library(devtools)

devtools::install_github("wrathematics/Rdym")
```

## Missingness
```r
NA
NA + 1
is.na(NA)
```


# Data Structures

* Vectors
* Factors
* Matrices
* Dataframes
* Some other things not worth getting into right now...

## Vectors
```{r}
c(1, 2, 3, 4, 5)
5:1
vec <- 1:5
vec[3] <- NA
vec
```

## Factors
```{r}
charvec <- c("a", "b", "a", "c", "a", "b", "b")
charvec
fac <- as.factor(charvec)
fac
```

## Matrices
```{r}
mat <- matrix(rnorm(21), nrow=7, ncol=3)
cov(mat)
```

## Dataframes
```{r}
data.frame(x=mat, charvec=charvec, fac=fac)
```


# Functions

* R is a (mostly) *functional* language.
* No macros, only functions.
* Powerful abstraction.

## Interacting with functions
```{r}
vec
median(vec)
median(na.rm=TRUE, vec)
median(vec, TRUE)
```

## Creating Functions
```{r}
vec

first_number_of_vec <- function(vec) vec[1]
first_number_of_vec(vec)

sum_of_first_and_last <- function(vec){
  sum <- vec[1] + vec[length(vec)]
  return(sum)
}
sum_of_first_and_last(vec)
```



## Questions?

This presentation is available at [github.com/wrathematics/2015SFSURworkshop](https://github.com/wrathematics/2015SFSURworkshop)

Exercises are also available there.