---
title: "Graphics the ggplot2 Way"
author: "Drew Schmidt"
date: "April 2, 2015"
output:
  ioslides_presentation:
    highlight: zenburn
    logo: pics/utk.png
    transition: faster
    smaller: no
    widescreen: yes
    css: style.css
---

## Contents



# Plotting Data in R

## Some Options

* Base graphics
* Grid
* Lattice
* ggplot2
* And [about 40 other packages](http://cran.r-project.org/web/views/Graphics.html)

```{r, echo=TRUE, eval=TRUE}
df <- data.frame(x=1:10, y=rnorm(10))
head(df)
```

## Base Graphics

Makes easy things easy, and hard things impossible.

```{r, echo=TRUE, eval=TRUE}
plot(df$x, df$y)
```


## Grid

"Assembly language" for graphics.



## Lattice

A powerful system for producing graphics, 

```{r, echo=TRUE, eval=TRUE}
library(lattice)
xyplot(y~x, df)
```

## ggplot2

Powerful and flexible; a darling of the R community.

```{r, echo=TRUE, eval=TRUE}
library(ggplot2)

ggplot(df, aes(x, y)) + geom_point()
```


# Introduction to ggplot2

## Ways of Plotting

* `qplot()` --- Shorthand
* Layer syntax --- Extremely rare "in the wild".
* geom/stat syntax --- Most popular method.
* autoplot --- Methods (as in OO) for plotting objects.


## General Procedure

* Put your data into a dataframe.
* Declare which dataframe ggplot should use and set aesthetics.
* Add layers via geom/stat functions.
* Set options, faceting, labels, titles, ...


## ?????

Geoms?  Aesthetics?  Facets?



# Deconstructing a ggplot2 Plot

## 4 Components

1. Geoms --- physical components (point, line, path, polygon)
2. Aesthetics --- visual cues (size, rotation, thickness, gradient, shape, color)
3. Coordinates
4. Faceting --- copolotting (more on this later)


## Customization Nearly Unlimited

![anatomy](./pics/ggplot2/ggplot2-anatomy-all-annotated.pdf)


## ggplot2

* Package by Hadley Wickham (2005)
* Implementation of Lee Wilkinson's Grammar of Graphics.
* Highly abstract.
* Powerful dsl for plotting data.






# How It Works Internally

## Creating the Plot

1. Construct
2. Build
3. Render
4. Draw


## Construct

* Expression is 
* Forces evaluation of formal ggplot2 command.


## Build

* All the "big decisions" are made and stored.
* Sets up scales, ranges, ... .


## Render

* Transforms data from build phase into a compact table.
* Collection of graphical objects (grobs).
    * Layout (rectangular, polar, ...)
    * Labels (x/y-axis, plot title, ...)
    * Axes


## Draw

* Uses grid graphics to produce the plot.


## All Together

```r
force(x)                     # Construct
data <- ggplot_build(x)      # Build
grob <- ggplot_gtable(data)  # Render
grid.draw(grob)              # Draw
```


# Example


## A Simple Scatterplot

```{r, echo=TRUE, eval=TRUE}
df <- data.frame(x=1:3, y=c(1.1, 2.1, 3.2))
df
```






